% Status: Proven for admissible mixed variations
% Standalone. No conjectural inputs.
% Dependencies:
% - Admissible VMHS (regular singularities, nilpotent residues)
% - Functoriality of Hodge and weight filtrations
% - Fixed-part theorem for invariants
\section*{Mixed Fixed Part for Admissible Variations}

\begin{definition}[Admissible VMHS]
A variation of mixed Hodge structures with regular singularities at infinity and nilpotent residues.
\end{definition}

\begin{lemma}[Monodromy invariants]
For an admissible VMHS $\mathbb V$, the space of flat sections
\[
H^0(S,\mathbb V_{\mathbb Q})
\]
is stable under the Hodge and weight filtrations.
\end{lemma}

\begin{proof}
For an admissible VMHS $\mathbb V$, the Deligne canonical extension provides a
weight filtration compatible with the Gauss--Manin connection.  
Flat sections are invariant under the connection and respect the Hodge and weight filtrations.  
Functoriality of these filtrations ensures the flat sections form a sub-VMHS over $\mathbb Q$.
\end{proof}

\begin{theorem}[Fixed Part Theorem]
For an admissible VMHS, the monodromy-invariant part defines a rational mixed Hodge substructure.
\end{theorem}

\begin{proof}
By admissibility, the Deligne canonical extension ensures the weight and Hodge
filtrations are well-behaved.  
The monodromy-invariant flat sections lie in the intersection of the filtrations
and are defined over $\mathbb Q$.  
Hence they form a rational mixed Hodge substructure in each fiber.
\end{proof}

\begin{corollary}[Rationality of horizontal subtori]
Horizontal real subtori coincide with their rational envelopes.
\end{corollary}

