\\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{microtype}
\usepackage{booktabs}

\title{Terminal Rigidity Witnesses and the Einstein--Rosen Bridge:\\
A Zero-Capacity Invariant for Physical Inference}
\author{Inacio F. Vasquez\\Independent Researcher}
\date{January 2026}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}
\newtheorem{axiom}{Axiom}

\newcommand{\Capacity}{\mathrm{Cap}}
\newcommand{\TC}{\mathrm{TC}}

\begin{document}
\maketitle

\begin{abstract}
We formalize the notion of a \emph{Terminal Rigidity Witness}: a physical system whose
global structure is nontrivial while its operational information capacity is exactly zero.
We show that the classical Einstein--Rosen bridge, arising as the maximal analytic extension
of the Schwarzschild solution, furnishes a canonical example. The result provides a sharp
stress test for claims of nonlocal physical inference and operational reconstruction.
\end{abstract}

\section{Transcript Capacity}

Let $X$ denote a hidden physical state and let $Y_{1:T}$ be the transcript produced by an
operational interface over $T$ steps. Define the \emph{transcript capacity} by
\[
\TC(T) := \sup I(X; Y_{1:T}),
\]
where the supremum ranges over all admissible encodings and observation strategies.
The asymptotic operational capacity of the system is
\[
\Capacity := \sup_{T \ge 1} \TC(T).
\]

\section{Operational Channels}

\begin{definition}
For a physical system $S$, the \emph{operational channel} $\mathcal C_S$ is the induced
input--output map from hidden states to observable transcripts available to any observer
respecting the causal and locality constraints of the underlying spacetime.
\end{definition}

\section{Terminal Rigidity}

\begin{definition}
A physical system $S$ is called a \emph{Terminal Rigidity Witness} if
\[
\Capacity(\mathcal C_S) = 0
\]
while the global structure of $S$ is nontrivial in the sense that it possesses
topological, geometric, or relational features not reducible to local operational data.
\end{definition}

The definition separates global structure from operational accessibility. A terminal
rigidity witness occupies a terminal element in the preorder of admissible inference
systems ordered by channel capacity.

\section{The Einstein--Rosen Bridge}

Consider the maximal analytic extension of the Schwarzschild spacetime, consisting of two
asymptotically flat regions joined by an Einstein--Rosen bridge.

\begin{axiom}[Topological Censorship]
No causal curve connects the two asymptotic regions of the maximally extended Schwarzschild
spacetime.
\end{axiom}

By topological censorship, any operational interface confined to causal propagation induces
a channel with no information flow between the two regions. Consequently, any transcript
available to an observer on one side is statistically independent of hidden states localized
on the other.

\begin{theorem}[Physical theorem; proof by topological censorship]
The Einstein--Rosen bridge is a terminal rigidity witness.
\end{theorem}

\begin{proof}
By topological censorship, the operational channel $\mathcal C_{\mathrm{ERB}}$ admits no
causal influence between the two asymptotic regions. Hence, for any transcript length $T$,
\[
I(X;Y_{1:T}) = 0,
\]
and therefore $\Capacity(\mathcal C_{\mathrm{ERB}})=0$. Nevertheless, the spacetime possesses
nontrivial global topology and geometry. Thus, the Einstein--Rosen bridge satisfies the
definition of a terminal rigidity witness.
\end{proof}

\section{Implications}

The Einstein--Rosen bridge provides a canonical example of a physical system with maximal
global structure and zero operational capacity. Any claim of nonlocal reconstruction,
information transfer, or operational geometry across the bridge must therefore violate
either locality, causality, or information accounting. Terminal rigidity witnesses thus
serve as invariants for ruling out spurious nonlocal inference mechanisms in both physics
and information-theoretic models.

\end{document}

