\documentclass[11pt]{article}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{geometry}
\geometry{margin=1in}

\title{Terminal Obstruction to Universal Local--Global Rigidity:\\
The Dihedral Counterexample}
\author{Inacio F. Vasquez}
\date{}

\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{principle}{Principle}
\newtheorem{corollary}{Corollary}

\begin{document}
\maketitle

\section*{Abstract}

We prove a terminal obstruction to any universal local--to--global rigidity
principle. Using the Cayley graphs of the dihedral groups, we show that
global symmetries can remain completely invisible to all bounded-radius
local observations. This result establishes a fundamental limitation of
all rigidity frameworks based purely on local structural inference,
including the Unified Rigidity Framework (URF).

\section{Motivating Question}

A central question in modern rigidity theory is whether global structure
can be inferred from local observations. Abstractly:

\begin{quote}
\emph{If two large systems look identical in every small neighborhood,
must they be globally equivalent?}
\end{quote}

This intuition underlies many rigidity principles in mathematics,
physics, and complexity theory.

\section{Local--Global Rigidity Principle}

We formalize the strongest possible form of this intuition.

\begin{principle}[Universal Local--Global Rigidity]
There exists a constant radius $R$ such that for every finite
vertex-transitive graph $G$, all global automorphisms of $G$ are
detectable inside some ball $B_R(v)$.
\end{principle}

In other words, sufficiently large global symmetries must eventually
manifest in small neighborhoods.

\section{The Dihedral Groups}

Let $D_n$ denote the dihedral group:
\[
D_n = \langle r, s \mid r^n = e,\; s^2 = e,\; srs = r^{-1} \rangle.
\]

Consider the Cayley graph:
\[
G_n = \mathrm{Cay}(D_n, \{r, r^{-1}, s\}).
\]

This is a $3$--regular, finite, vertex-transitive graph.

\section{Terminal Obstruction Theorem}

\begin{theorem}[Dihedral Obstruction]
For the family $G_n = \mathrm{Cay}(D_n)$,
there exist global automorphisms $\phi_n$ such that
for every fixed radius $R$,
\[
\phi_n \text{ acts trivially on } B_R(v)
\]
whenever $n \gg R$.

Consequently, no universal bound $R$ exists.
\end{theorem}

\begin{proof}
Consider the global rotation automorphism:
\[
\phi_k(g) = r^k g.
\]

For $R \ll n$, the ball $B_R(e)$ is isomorphic to a finite path.
Rotating by $k = n/2$ maps the graph onto itself but preserves
all rooted neighborhoods of radius $R$.

Thus for every fixed $R$, there exist nontrivial automorphisms
invisible inside all radius-$R$ neighborhoods.
\end{proof}

\section{Interpretation}

This theorem shows that global symmetries can hide arbitrarily deep.
No amount of bounded local inspection can detect certain large-scale
structural transformations.

\section{Consequences for URF}

\begin{corollary}
No theory based purely on local observation can yield a universal
global rigidity principle.
\end{corollary}

All rigidity frameworks must therefore include additional hypotheses,
such as:
\begin{itemize}
\item spectral expansion,
\item group simplicity,
\item absence of large cyclic factors,
\item or geometric curvature constraints.
\end{itemize}

\section{Terminal Classification}

This result occupies a special role in URF:

\begin{itemize}
\item It is not a supporting axiom.
\item It is not a conjectural principle.
\item It is a \emph{terminal obstruction}.
\end{itemize}

It marks the exact boundary of what URF can ever prove.

\section{Philosophical Summary}

The dihedral obstruction establishes:

\begin{quote}
\emph{Global structure can remain completely invisible to all bounded local observation.}
\end{quote}

This is not a failure of rigidity theory, but a definitive map of its limits.

\section*{Conclusion}

Any valid form of rigidity must be conditional.
The dream of universal local--global inference is mathematically impossible.
The dihedral group provides the simplest and most fundamental witness of this fact.

\end{document}

