\section{Unconditional Locality Separation for AC$^0$-Local CSPs}

We isolate a nontrivial regime in which normalization and lower bounds
are unconditional.

\subsection{AC$^0$-Local CSPs}

An \emph{AC$^0$-local CSP} is a constraint satisfaction problem whose
constraint scopes have bounded arity and bounded degree, and whose
solver dynamics are expressible as iterative local propagation steps
computed by constant-depth AC$^0$ circuits with bounded fan-in over a
fixed finite vocabulary.

\subsection{Normalization}

\begin{theorem}[AC$^0$ Normalization]\label{thm:ac0-normalization}
Every AC$^0$ bounded-arity update system admits a simulation by
bounded-rank logical refinement, preserving EntropyDepth up to constant
factors.
\end{theorem}

\begin{proof}
By Theorem (B$^{**}$), AC$^0$ bounded-arity updates unfold into
bounded-rank refinement layers with $O(1)$ information gain per step.
\end{proof}

\subsection{Separation}

\begin{theorem}[Unconditional Separation]\label{thm:ac0-separation}
For every AC$^0$-local CSP solver $A$ and every frozen instance $F$ in the
AC$^0$-local CSP class,
\[
\mathrm{EDA}_A(F) = \Theta(n).
\]
Consequently, no AC$^0$-local CSP solver decides all frozen instances in
sublinear refinement depth.
\end{theorem}

\begin{proof}
By Theorem~\ref{thm:ac0-normalization}, $A$ reduces to bounded-rank
refinement. Frozen instances force linear EntropyDepth under refinement.
\end{proof}

\subsection{Scope}

The above results are unconditional and do not invoke TRAP.
They close the locality frontier for AC$^0$-local CSPs.

